Models

### 2023-03-02 Code/Models.Rmd
Not sure what I changed but small

```{r}
old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
```

@@

paste("Linear without year effect has AIC of", AIC(men_finals_one), "and log-lik of", logLik(men_finals_one))
paste("Linear with year effect has AIC of", AIC(men_finals_year), "and log-lik of", logLik(men_finals_year))
paste("Gamma without year effect has AIC of", AIC(men_finals_one_gamma), "and log-lik of", logLik(men_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_finals_year_gamma), "and log-lik of", logLik(men_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_finals_one_gamma_log), "and log-lik of", logLik(men_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_finals_year_gamma_log), "and log-lik of", logLik(men_finals_year_gamma_log))
```
```{}
The best selected model is men_finals_year_gamma as that has the smallest AIC: -491 and largest log-lik of 248.5
```

```{r}
library(lme4)
old_finals_one <- glm(RxnTime ~ 1, data = old_finals)
old_finals_year <- lmer(RxnTime ~ (1|Year), data = old_finals)
old_finals_one_gamma <- glm(RxnTime ~ 1, data = old_finals, family = Gamma)
old_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = old_finals, family = Gamma)
old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = old_finals, family = Gamma(link = "log"))
old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = old_finals, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(old_finals_one), "and log-lik of", logLik(old_finals_one))
paste("Linear with year effect has AIC of", AIC(old_finals_year), "and log-lik of", logLik(old_finals_year))
paste("Gamma without year effect has AIC of", AIC(old_finals_one_gamma), "and log-lik of", logLik(old_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(old_finals_year_gamma), "and log-lik of", logLik(old_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(old_finals_one_gamma_log), "and log-lik of", logLik(old_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(old_finals_year_gamma_log), "and log-lik of", logLik(old_finals_year_gamma_log))
```
```{}
The best selected model is old_finals_year_gamma as that has the smallest AIC: -465 and largest log-lik of 235.5.  The models for the older data does not seem to fit the data as good as for the full data set.
```

@@

```{r}
library(lme4)
old_pooled_one <- glm(RxnTime ~ 1, data = old_pooled)
old_pooled_year <- lmer(RxnTime ~ (1|Year), data = old_pooled)
old_pooled_one_gamma <- glm(RxnTime ~ 1, data = old_pooled, family = Gamma)
old_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = old_pooled, family = Gamma)
old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = old_pooled, family = Gamma(link = "log"))
old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = old_pooled, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(old_pooled_one), "and log-lik of", logLik(old_pooled_one))
paste("Linear with year effect has AIC of", AIC(old_pooled_year), "and log-lik of", logLik(old_pooled_year))
paste("Gamma without year effect has AIC of", AIC(old_pooled_one_gamma), "and log-lik of", logLik(old_pooled_one_gamma))
paste("Gamma with year effect has AIC of", AIC(old_pooled_year_gamma), "and log-lik of", logLik(old_pooled_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(old_pooled_one_gamma_log), "and log-lik of", logLik(old_pooled_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(old_pooled_year_gamma_log), "and log-lik of", logLik(old_pooled_year_gamma_log))
```
```{}
The best model is old_pooled_year_gamma as that has the smallest AIC: -1687.5 and the largest log-lik: 846.8
```

@@

```{r}
set.seed(1)
simulation <- function(model, base){
  simfit <- function(model, n = 12000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(mean(x < 0.1))
  print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
}
simulation(men_finals_year_gamma_log, men_finals_one_gamma_log)
```

@@

```{r}
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "ds")
clusWilcox.test(Time, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "rgl")
```

@@

```{r}
In both examples, rgl produced a smaller p value
```

>>>>

```{r}
men_old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
men_old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
```

@@

paste("Linear without year effect has AIC of", AIC(men_finals_one), "and log-lik of", logLik(men_finals_one))
paste("Linear with year effect has AIC of", AIC(men_finals_year), "and log-lik of", logLik(men_finals_year))
paste("Gamma without year effect has AIC of", AIC(men_finals_one_gamma), "and log-lik of", logLik(men_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_finals_year_gamma), "and log-lik of", logLik(men_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_finals_one_gamma_log), "and log-lik of", logLik(men_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_finals_year_gamma_log), "and log-lik of", logLik(men_finals_year_gamma_log))
```

```{}
The best selected model is men_finals_year_gamma as that has the smallest AIC: -491 and largest log-lik of 248.5
```

```{r}
library(lme4)
men_old_finals_one <- glm(RxnTime ~ 1, data = men_old_finals)
men_old_finals_year <- lmer(RxnTime ~ (1|Year), data = men_old_finals)
men_old_finals_one_gamma <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma)
men_old_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma)
men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma(link = "log"))
men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(men_old_finals_one), "and log-lik of", logLik(men_old_finals_one))
paste("Linear with year effect has AIC of", AIC(men_old_finals_year), "and log-lik of", logLik(men_old_finals_year))
paste("Gamma without year effect has AIC of", AIC(men_old_finals_one_gamma), "and log-lik of", logLik(men_old_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_old_finals_year_gamma), "and log-lik of", logLik(men_old_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_old_finals_one_gamma_log), "and log-lik of", logLik(men_old_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_old_finals_year_gamma_log), "and log-lik of", logLik(men_old_finals_year_gamma_log))
```
```{}
The best selected model is men_old_finals_year_gamma as that has the smallest AIC: -465 and largest log-lik of 235.5.  The models for the older data does not seem to fit the data as good as for the full data set.
```

@@

```{r}
library(lme4)
men_old_pooled_one <- glm(RxnTime ~ 1, data = men_old_pooled)
men_old_pooled_year <- lmer(RxnTime ~ (1|Year), data = men_old_pooled)
men_old_pooled_one_gamma <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma)
men_old_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma)
men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma(link = "log"))
men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(men_old_pooled_one), "and log-lik of", logLik(men_old_pooled_one))
paste("Linear with year effect has AIC of", AIC(men_old_pooled_year), "and log-lik of", logLik(men_old_pooled_year))
paste("Gamma without year effect has AIC of", AIC(men_old_pooled_one_gamma), "and log-lik of", logLik(men_old_pooled_one_gamma))
paste("Gamma with year effect has AIC of", AIC(men_old_pooled_year_gamma), "and log-lik of", logLik(men_old_pooled_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(men_old_pooled_one_gamma_log), "and log-lik of", logLik(men_old_pooled_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(men_old_pooled_year_gamma_log), "and log-lik of", logLik(men_old_pooled_year_gamma_log))
```
```{}
The best model is men_old_pooled_year_gamma as that has the smallest AIC: -1687.5 and the largest log-lik: 846.8
```

@@

```{r}
set.seed(1)
simulation <- function(model, base){
  simfit <- function(model, n = 12000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(mean(x < 0.1))
  print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
}
simulation(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log)
```

@@

```{r}
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "ds")
clusWilcox.test(Time, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "rgl")
#pairwise.clusWilcox.test(comp$Time, group= comp$Year, cluster = comp$Athlete, alternative = "greater", method = "rgl")
```

@@

```{}
In both examples, rgl produced a smaller p value
```

```{r}
mini <- min(aggregate(times$RxnTime, list(times$Batch), FUN =mean)$x)
maxi <- max(aggregate(times$RxnTime, list(times$Batch), FUN =mean)$x)
cat(maxi - mini)
```
